Re-analysing the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}): A textbook illustration of the absence of evidence fallacy

### 2020-12-16 manuscript.tex
updating the manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man, donotrepeattitle,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdflang={en-EN},
  pdfkeywords={NHST, Bayesian, logical fallacy, reanalysis, inner speech, rumination, electromyography},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{The absence of evidence fallacy}
\author{Ladislas Nalborczyk\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Univ. Grenoble Alpes, CNRS, Grenoble INP, GIPSA-lab, 38000 Grenoble, France}
\authornote{

Correspondence concerning this article should be addressed to Ladislas Nalborczyk, GIPSA-lab, CNRS, Univ. Grenoble Alpes, 11 Rue des Mathématiques, 38400 Saint-Martin-d'Hères, France. E-mail: ladislas.nalborczyk@gipsa-lab.fr}
\keywords{NHST, Bayesian, logical fallacy, reanalysis, inner speech, rumination, electromyography}
\usepackage{lineno}

@@

\title{\textbf{Submitted to Meta-Psychology. Click here to follow the fully transparent editorial process of this submission. Participate in open peer review by commenting through hypothes.is directly on this preprint.}\\
~\\
~\\
Re-analysing the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}): A textbook illustration of the absence of evidence fallacy}

\date{}

@@

Wordcount (excluding abstract, references, tables, and figures): 4579

@@

To model EMG peak amplitude variations in response to the rumination and distraction inductions, we fitted a Bayesian multivariate regression model with the natural logarithm of the EMG peak amplitude as an outcome and \emph{Condition} (baseline, rumination, distraction) as a categorical predictor. Therefore, the intercept represents the estimated natural logarithm of the EMG peak amplitude in the baseline condition, and the slopes for the rumination and distraction conditions represent deviations from the baseline. These analyses were conducted using the \texttt{brms} package (Bürkner, \protect\hyperlink{ref-R-brms}{2017}), an \texttt{R} implementation of Bayesian multilevel models that employs the probabilistic programming language \texttt{Stan} (Carpenter et al., \protect\hyperlink{ref-carpenter_stan_2017}{2017}). We ran four chains including each 10.000 iterations and a warmup of 2.000 iterations. Posterior convergence was assessed examining autocorrelation and trace plots, as well as the Gelman-Rubin statistic. Constant effects estimates were summarised via their posterior mean and 95\% credible interval. We also report Bayes factors (BFs) computed using the Savage-Dickey method.\footnote{This method consists in taking the ratio of the posterior density at the point of interest divided by the prior density at that point (Wagenmakers et al., \protect\hyperlink{ref-wagenmakers_bayesian_2010}{2010}).} These BFs can be interpreted as updating factors, from prior knowledge (what we knew before seeing the data) to posterior knowledge (what we know after seeing the data). A summary of the estimations from this model is presented in Table \ref{tab:summary}. This analysis revealed strong evidence for the hypothesis of a higher average EMG peak amplitude in the rumination condition as compared to the baseline condition for both the FRO and OOI muscles (as assessed by the BFs). However, the BFs supported the null hypothesis (i.e., the hypothesis of no difference) between the baseline and distraction conditions for the FRO and were inconclusive for both the OOI and OOS muscles.

@@

This problem is tackled in modern usages of null-hypothesis significance tests by ensuring that the claim under scrutiny is submitted to \emph{severe} tests (e.g., Mayo \& Spanos, \protect\hyperlink{ref-mayo_severe_2006}{2006}; Mayo, \protect\hyperlink{ref-mayo_statistical_2018}{2018}). In general terms, the strong severity principle states that we have evidence for a claim to the extent that it survives a stringent scrutiny, that is, to the extent that it survives \emph{severe} tests. More precisely, some claim (e.g., \(\theta = 0\)) is said to be \emph{severely tested} if it had great chances of being falsified, had the claim been false. When a statistical test is under-powered (for detecting a given effect size), the claim under scrutiny is not strongly (severely) tested, hence it not possible to obtain strong or reliable evidence for the claim (bad test, no evidence).

@@

We suggest the (a priori) power of the study ran by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter (which seems reasonable given the high inter-individual variability in such effects, cf.~the next section but also Nalborczyk, Grandchamp, et al., \protect\hyperlink{ref-nalborczyk_can_2020}{2020}), therefore the a priori power of the main statistical test from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was around \(0.42\), meaning that they had less than 1 chance out of 2 to find a significant effect (given that the population effect size was actually \(0.36\)).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# A priori power for n = 26 and d = 0.36}
\KeywordTok{library}\NormalTok{(pwr)}
\KeywordTok{pwr.t.test}\NormalTok{(}
  \DataTypeTok{n =} \DecValTok{26}\NormalTok{, }\DataTypeTok{d =} \FloatTok{0.72} \OperatorTok{/}\StringTok{ }\DecValTok{2}\NormalTok{, }\DataTypeTok{sig.level =} \FloatTok{0.05}\NormalTok{,}
  \DataTypeTok{type =} \StringTok{"one.sample"}\NormalTok{, }\DataTypeTok{alternative =} \StringTok{"two.sided"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 
##      One-sample t test power calculation 
## 
##               n = 26
##               d = 0.36
##       sig.level = 0.05
##           power = 0.4228455
##     alternative = two.sided
\end{verbatim}

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the \emph{relative evidence} for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we present below the results of two simulations.

@@

First, we simulated 10.000 datasets (for \(N = 26\)) under the assumption of either no effect (i.e., the null hypothesis of \(d = 0\)), an effect size of \(d = 0.36\) (i.e., the supposed target effect size in Moffatt et al., \protect\hyperlink{ref-moffatt_inner_2020}{2020}), or an effect size of \(d = 0.72\) (i.e., the effect size reported in Nalborczyk et al., \protect\hyperlink{ref-nalborczyk_orofacial_2017}{2017}). As shown in Figure \ref{fig:bf-dance}, the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, 6.46\% of the computed BFs are above 0 and hence support the alternative hypothesis (although the ``true'' effect size is \(d = 0\)). When the ``true'' effect size is of \(d = 0.36\), 72.06\% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the ``true'' effect size is of \(d = 0.72\), 24.46\% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates.\footnote{To assess the extent to which the BF computed for the OOI by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) (i.e., \(\text{BF}_{10} = 0.278\)) is ``surprising'' given or ``compatible'' with the hypothesis of an effect size of \(d = 0.36\), we can compute the probability of obtaining this finding or a more extreme finding given the hypothesis, which is approximately equal to 0.29.}

Second, we conducted another simulation with the aim of assessing the relation between the sample size and the value of the BF. In the previous section, we fitted a multivariate Bayesian regression model with varying-intercepts by participant and weakly informative priors on the EMG data collected by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}). Using this model, we i) generated new datasets from the posterior predictive distribution of this model and ii) we computed the BF in favour of the alternative hypothesis (\(\text{BF}_{10}\)) using the \texttt{BayesFactor} package (Morey \& Rouder, \protect\hyperlink{ref-R-BayesFactor}{2018}). We used a ``medium'' prior (i.e., \(r = \sqrt{2}/2\)) on the scale parameter of the Cauchy prior for the alternative hypothesis. We repeated this procedure for varying sample sizes from 20 to 200 participants (by increments of 10 participants) with \(1000\) simulations (i.e., 1000 simulated datasets) for each sample size.

{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/simulated-power-1}

\caption{Average natural logarithm of the Bayes factor in favour of the alternative hypothesis, along with its standard error, computed over 1000 datasets of increasing size simulated from the posterior predictive distribution of the varying-intercept multivariate Bayesian regresion model, fitted on the data from Moffatt et al. (2020). A log-BF belows 0 represents evidence for the null hypothesis (relative to the alternative hypothesis) and a log-BF above 0 represents evidence for the alternative hypothesis (relative to the null hypothesis).}\label{fig:simulated-power}
\end{figure}

As shown in Figure \ref{fig:simulated-power}, the natural logarithm of the BF in favour of the alternative hypothesis is growing proportionally with sample size. More precisely, whereas BFs computed on small samples (i.e., below 80 participants) support the null hypothesis, BFs computed on larger samples support the alternative hypothesis for all three facial muscles. For instance, the average \(\text{BF}_{10}\) computed for the OOI muscle with a sample size of 160 participants is of \(\exp(2.18) \approx 8.85\), indicating that these data are approximately \(8.85\) times more likely under the alternative hypothesis than under the null hypothesis. To sum up, this reveals that although at low sample sizes, the BF may provide (weak) evidence for the null hypothesis (relative to the alternative hypothesis), this pattern may very well reverse for higher sample sizes.

We should keep in mind some limitations of this analysis, which uses simulated datasets form the posterior predictive distribution estimated on the data collected by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}). This analysis resembles to the Bayesian analogue of the frequentist post-hoc power analysis, which has been much criticised (e.g., Lakens, \protect\hyperlink{ref-lakens_20_2014}{2014}). A crucial assumption of the present analysis is that the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) is our best source of information regarding the main effect of interest. However, the present analysis also differs from the frequentist post-hoc power analysis on several grounds. First, with the present analysis, we do not aim to assess the ability of our statistical test to pass some dichotomic threshold (e.g., accept/reject). Instead, we aim to assess how the \(\text{BF}_{10}\) (i.e., the evidence for the alternative hypothesis, relative to the null hypothesis) behaves with varying sample sizes. Second, the present analysis relies on the posterior predictive distribution of the model fitted on the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}), which naturally incorporates uncertainty about the effect of interest. By simulating datasets of varying sample sizes from the posterior predictive distribution (and by relying on a large number of simulations), uncertainty about the effect size is naturally incorporated into the results of the simulation.

@@

\leavevmode\hypertarget{ref-lakens_20_2014}{}%
Lakens, D. (2014). The 20\% Statistician: Observed power, and what to do if your editor asks for post-hoc power analyses. In \emph{The 20\% Statistician}.

@@

\leavevmode\hypertarget{ref-R-BayesFactor}{}%
Morey, R. D., \& Rouder, J. N. (2018). \emph{BayesFactor: Computation of bayes factors for common designs}. \url{https://CRAN.R-project.org/package=BayesFactor}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man, donotrepeattitle,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Re-analysing the data from Moffatt et al. (2020): A textbook illustration of the absence of evidence fallacy},
  pdfauthor={Ladislas Nalborczyk1},
  pdflang={en-EN},
  pdfkeywords={NHST, Bayesian, logical fallacy, reanalysis, inner speech, rumination, electromyography},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{The absence of evidence fallacy}
\keywords{NHST, Bayesian, logical fallacy, reanalysis, inner speech, rumination, electromyography}
\usepackage{lineno}

@@

\title{Re-analysing the data from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}): A textbook illustration of the absence of evidence fallacy}
\author{Ladislas Nalborczyk\textsuperscript{1}}
\date{}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Univ. Grenoble Alpes, CNRS, Grenoble INP, GIPSA-lab, 38000 Grenoble, France}

@@

Wordcount (excluding abstract, references, tables, and figures): 4182

@@

To model EMG peak amplitude variations in response to the rumination and distraction inductions, we fitted a Bayesian multivariate regression model with the natural logarithm of the EMG peak amplitude as an outcome and \emph{Condition} (baseline, rumination, distraction) as a categorical predictor. Therefore, the intercept represents the estimated natural logarithm of the EMG peak amplitude in the baseline condition, and the slopes for the rumination and distraction conditions represent deviations from the baseline. These analyses were conducted using the \texttt{brms} package (Bürkner, \protect\hyperlink{ref-R-brms}{2017}), an \texttt{R} implementation of Bayesian multilevel models that employs the probabilistic programming language \texttt{Stan} (Carpenter et al., \protect\hyperlink{ref-carpenter_stan_2017}{2017}). We ran four chains including each 10.000 iterations and a warmup of 2.000 iterations, using weakly informative priors on all parameters (for more details, see the \protect\hyperlink{supp}{supplementary materials}). Posterior convergence was assessed examining autocorrelation and trace plots, as well as the Gelman-Rubin statistic. Constant effects estimates were summarised via their posterior mean and 95\% credible interval. We also report Bayes factors (BFs) computed using the Savage-Dickey method.\footnote{This method consists in taking the ratio of the posterior density at the point of interest divided by the prior density at that point (Wagenmakers et al., \protect\hyperlink{ref-wagenmakers_bayesian_2010}{2010}).} These BFs can be interpreted as updating factors, from prior knowledge (what we knew before seeing the data) to posterior knowledge (what we know after seeing the data). A summary of the estimations from this model is presented in Table \ref{tab:summary}. This analysis revealed strong evidence for the hypothesis of a higher average EMG peak amplitude in the rumination condition as compared to the baseline condition for both the FRO and OOI muscles (as assessed by the BFs). However, the BFs supported the null hypothesis (i.e., the hypothesis of no difference) between the baseline and distraction conditions for the FRO and were inconclusive for both the OOI and OOS muscles.

@@

This problem is tackled in modern usages of null-hypothesis significance tests by ensuring that the claim under scrutiny is submitted to \emph{severe} tests (e.g., Mayo \& Spanos, \protect\hyperlink{ref-mayo_severe_2006}{2006}; Mayo, \protect\hyperlink{ref-mayo_statistical_2018}{2018}). In general terms, the strong severity principle states that we have evidence for a claim to the extent that it survives a stringent scrutiny, that is, to the extent that it survives \emph{severe} tests. More precisely, some claim (e.g., \(\theta = 0\)) is said to be \emph{severely tested} if it had great chances of being corroborated/falsified, had the claim been true/false. When a statistical test is under-powered (for detecting a given effect size), the claim under scrutiny is not strongly (severely) tested, hence it not possible to obtain strong or reliable evidence for the claim (bad test, no evidence).

@@

{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/power-1}

\caption{Statistical power as a function of both sample size and effect size, for a one-sample t-test with a significance level of 0.05. The white dot indicates the minimal effect size that can be detected with a probability equal or superior to 0.9 with a sample size of N = 26.}\label{fig:power}
\end{figure}

We suggest the (a priori) power of the study ran by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was much lower than suggested by the authors. Indeed, we speculate that the standardised mean difference in EMG peak amplitude between the rumination and distraction conditions may be much weaker than the standardised mean difference in EMG amplitude between the rumination and rest conditions. If we assume that the former is half the size of the latter (which seems reasonable given the high inter-individual variability in such effects, cf.~the next section but also Nalborczyk, Grandchamp, et al., \protect\hyperlink{ref-nalborczyk_can_2020}{2020}), therefore the a priori power of the main statistical test from Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) was around \(0.42\), meaning that they had less than 1 chance out of 2 to find a significant effect (given that the population effect size was actually \(0.36\)). Notice that whereas taking half the effect size of Nalborczyk et al. (\protect\hyperlink{ref-nalborczyk_orofacial_2017}{2017}) may seem arbitrary, Figure \ref{fig:power} shows that a one-sample t-test with a sample size of \(N = 26\) is under-powered for a vast range of effect sizes.

@@

While we appreciate the effort, the current approach poses new problems. First, contrary to what the authors suggest, whereas computing a BF indeed allows assessing the \emph{relative evidence} for the null, computing a BF (i.e., comparing two models) does not solve at all the problem of low power. More precisely, the sensitivity (i.e., the ability to attain a certain goal) of an experimental design to detect a given effect is an issue for both frequentist and Bayesian statistical tests. To illustrate this point, we simulated 10.000 datasets (for \(N = 26\)) under the assumption of either no effect (i.e., the null hypothesis of \(d = 0\)), an effect size of \(d = 0.36\) (i.e., the supposed target effect size in Moffatt et al., \protect\hyperlink{ref-moffatt_inner_2020}{2020}), or an effect size of \(d = 0.72\) (i.e., the effect size reported in Nalborczyk et al., \protect\hyperlink{ref-nalborczyk_orofacial_2017}{2017}).

@@

As shown in Figure \ref{fig:bf-dance}, the distribution of BFs computed under each hypothesis reveals important inter-simulation variability. For instance, under the null hypothesis, 6.46\% of the computed BFs are above 0 and hence support the alternative hypothesis (although the ``true'' effect size is \(d = 0\)). When the ``true'' effect size is of \(d = 0.36\), 72.06\% of the BFs are below 0 and hence support the null hypothesis (although the true effect size is actually non-null). When the ``true'' effect size is of \(d = 0.72\), 24.46\% of the BFs are still below 0. In other words, for small sample and effect sizes, BFs have high error rates. To assess the extent to which the BF computed for the OOI by Moffatt et al. (\protect\hyperlink{ref-moffatt_inner_2020}{2020}) (i.e., \(\text{BF}_{10} = 0.278\)) is ``surprising'' given or ``compatible'' with some effect size, we can compute the proportion of simulated BFs that are equal or more extreme than the one reported by the authors for each hypothesis, which is approximately equal to 0.57, 0.29, and 0.03, for the hypothesis of \(d = 0\), \(d = 0.36\), and \(d = 0.72\), respectively. In other words, observing a \(\text{BF}_{10} = 0.278\) is more compatible with the hypothesis of \(d = 0\) (i.e., the null hypothesis) than with the hypothesis of \(d = 0.36\) or \(d = 0.72\) (at this sample size).

@@

@@
